We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-01-24 paper/paper.tex
edit abstract

\begin{abstract}
We quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use simulations to build an explicit probability distribution for the
asphericity around galaxies similar to the LG in the Lambda Cold Dark
Matter paradigm.
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS. 
We find as a common trends in thse simulations that: i) at most of
$1\%$ the pairs are expected  to have satellite distributions with the
same degree of deviation from sphericity as the Local Group; ii) between
$38\%$ to $59\%$ of the pairs have a halo with a satellite
distribution as aspherical as  in M31; and iii) at most $2\%$ have a
satellite distribution as planar as in the MW. 
These results place the LG close to the level of a 3-$\sigma$ outlier
in context provided by LCDM simulations.
\end{abstract}

\begin{keywords}Galaxies: halos --- Galaxies: high-redshift --- Galaxies: statistics
--- Dark Matter --- Methods: numerical 
\end{keywords}

>>>>

\begin{abstract}
We quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use simulations to build an explicit probability distribution for the
asphericity around galaxies similar to the LG in the Lambda Cold Dark
Matter (LCDM) paradigm.
This allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that fully resemble the LG.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS. 
We find three common trends in these simulations. First, at most of
$1\%$ the pairs are expected to have satellite distributions with the
same degree of deviation from sphericity as the Local Group; second, between
$38\%$ to $59\%$ of the pairs have a halo with a satellite
distribution as aspherical as  in M31; and third, at most $2\%$ of the
pairs have a satellite distribution as planar as in the MW. 
The M31 galaxy turns out to have a rather typical distribution 
in the LCDM context, for its brightest satellites, while the MW is
atypical in this respect. 
These results place the LG close to the level of a 3-$\sigma$ outlier
in the LCDM paradigm.
We summarize our results in terms of mean values and covariance
matrices. This approach will facilitate the direct comparison of the
results from different numerical setups and choices to select LG pairs
in simulations. 
\end{abstract}

\begin{keywords}Galaxies: halos --- Galaxies: statistics --- Dark
  Matter --- Methods: numerical  
\end{keywords}
